---
title: "Bayes Theorem"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```

```{r}
library(tidyverse)
library(ggthemes)
```

## Formula

$P(A \mid B) = \frac{P(B \mid A) \, P(A)}{P(B)}$


$P(Theory \mid Data) = \frac{P(Data \mid Theory) \, P(Theory)}{P(Data)}$

$Posterior: P(Theory \mid Data) \quad Likelihood: P(Data \mid Theory) \quad Prior: P(Theory)  \quad Evidenz:  {P(Data)}$





### Examples

Diagnostics of AIDS

1 out of thousands has AIDS.


Test:

Sensitivity: P(+|AIDS) = 0.98

Specificity: P(-|noAIDS) = 0.97

$P(+|AIDS) = 0.98 \quad  P(-|AIDS) = 0.02 \quad P(+|noAIDS) = 0.03  \quad P(-|noAIDS) = 0.97$



$P(AIDS\mid+) = \frac{P(+\mid AIDS) \, P(AIDS)}{P(+\mid AIDS)P(AIDS) + P(+\mid noAIDS)P(noAIDS)} = \frac{0.98\times0.001}{0.98\times0.001 + 0.03\times0.999} = 0.032$




```{r}
`p(a|b)` <- function(`p(b|a)`, `p(a)`, `p(b)`) {
  `p(b|a)` * `p(a)` / `p(b)`
}

`p(b)` <- function(`p(a)`, `p(b|a)`, `p(¬a)`, `p(b|¬a)`) {
  `p(a)` * `p(b|a)` + `p(¬a)` *  `p(b|¬a)`
}


calc_prob_of_positive <-
  function(infection_rate, sensitivity, specificity) {
    `p(a|b)`(
      sensitivity,
      infection_rate,
      `p(b)`(
        sensitivity,
        infection_rate,
        1 - specificity,
        1 - infection_rate
      )
    )
  }

calc_prob_of_negative <-
  function(infection_rate, sensitivity, specificity) {
    `p(a|b)`(
      specificity,
      1-infection_rate,
      `p(b)`(
        specificity,
        1-infection_rate,
        1-sensitivity,
        infection_rate
      )
    )
  }

```

```{r}

data <-
  expand_grid(
    infection_rate = seq(0.005, 0.99, .005),
    sensitivity = seq(0.9, 0.99, 0.001),
    specificity = seq(0.9, 0.99, 0.001)
  ) %>%
  mutate(probability_positive = calc_prob_of_positive(infection_rate, sensitivity, specificity))
```

```{r fig.height=30, fig.width=30}
data %>% 
  mutate(infection_rate_bin = cut(infection_rate,10), 
         probability_positive_quality=ifelse(probability_positive > 0.9, "Good","Bad")) %>% 
  ggplot(aes(x=sensitivity, y=specificity, color=probability_positive_quality)) +
  geom_point(size=4) +
  scale_color_manual(values = c("Bad" = "#E15759", "Good" = "#59A14F")) +
  facet_wrap(~infection_rate_bin)
```

```{r}
read_csv("countries_infection_rates.csv")
```


```{r}
countries_infection_rates <- expand_grid(
  read_csv("countries_infection_rates.csv") %>% 
  mutate_if(is.numeric, funs((.)/100)),
  sensitivity = seq(0.9, 0.99, 0.001),
  specificity = seq(0.9, 0.99, 0.001)
)  %>% 
  mutate(probability_positive= calc_prob_of_positive(infection_rate, sensitivity, specificity)) %>% 
  mutate(probability_positive_quality=ifelse(probability_positive > 0.6, "Good","Bad"))
  
```

```{r fig.height=20, fig.width=20}
countries_infection_rates %>% 
  ggplot(aes(x=sensitivity, y=specificity, color=probability_positive_quality)) +
  geom_point() + 
  scale_color_manual(values = c("Bad" = "#E15759", "Good" = "#59A14F")) +
  facet_wrap(~country)
  
  

```



