---
title: "Bayes Theorem"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```

```{r}
library(tidyverse)
```

## Formula

$P(A \mid B) = \frac{P(B \mid A) \, P(A)}{P(B)}$

$P(Theory \mid Data) = \frac{P(Data \mid Theory) \, P(Theory)}{P(Data)}$

$Posterior: P(Theory \mid Data) \quad Likelihood: P(Data \mid Theory) \quad Prior: P(Theory)  \quad Evidenz:  {P(Data)}$





### Examples

Diagnostics of AIDS

1 out of thousands has AIDS.


Test:

Sensitivity: P(+|AIDS) = 0.98

Specificity: P(-|noAIDS) = 0.97

$P(+|AIDS) = 0.98 \quad  P(-|AIDS) = 0.02 \quad P(+|noAIDS) = 0.03  \quad P(-|noAIDS) = 0.97$



$P(AIDS\mid+) = \frac{P(+\mid AIDS) \, P(AIDS)}{P(+\mid AIDS)P(AIDS) + P(+\mid noAIDS)P(noAIDS)} = \frac{0.98\times0.001}{0.98\times0.001 + 0.03\times0.999} = 0.032$


```{r}
```


```{r}


calc_prob_of_positive <-
  function(infection_rate, sensitivity, specificity) {
    sensitivity * infection_rate / (sensitivity * infection_rate + (1 - specificity) * (1 -
                                                                                          infection_rate))
  }

data <-
  expand_grid(
    infection_rate = seq(0.01, 0.96, .05),
    sensitivity = seq(0.9, 0.99, 0.01),
    specificity = seq(0.9, 0.99, 0.01)
  ) %>%
  mutate(probability_positive = calc_prob_of_positive(infection_rate, sensitivity, specificity))
```

```{r}
data %>% arrange(probability_positive) 
```

